---
title: "01: Data cleaning and decisions"
---

## Purpose

This document shows where data cleaning happens and, just as importantly, where the student explains what was changed and why. In a thesis repository, this is often the most important analytical chapter because it preserves the chain of decisions.

## Recommended raw files

Store raw data in `data/` and avoid editing those files directly. Instead, read them in and create cleaned objects in code.

```{r}
#| eval: false
library(tidyverse)

surveys_raw <- read_csv("data/surveys.csv")
phenology_raw <- read_csv("data/phenology.csv")
spatial_raw <- read_csv("data/spatial_sites.csv")
```

## Example cleaning workflow

```{r}
#| eval: false
surveys_df <- surveys_raw |>
  janitor::clean_names() |>
  mutate(
    date = as.Date(date),
    habitat_type = factor(habitat_type),
    caste = factor(caste),
    survey_id = as.character(survey_id)
  ) |>
  filter(!is.na(site_id), !is.na(date))
```

## Validation checks

The repository includes a validation function in `R/analysis_functions.R`.

```{r}
#| eval: false
source("R/analysis_functions.R")
validation_results <- validate_survey_data(surveys_df)
validation_results$missing
validation_results$impossible
validation_results$duplicates
```

## Decisions to document

In a real project, students should explain decisions such as:

- how dates were standardized,
- how missing values were handled,
- which rows were excluded,
- how factor levels were defined,
- and whether any derived variables were created.

## Good practice

Write cleaning as a narrative record, not just a code dump. A committee member should be able to understand what changed without reverse-engineering every line.
