---
title: "02: Exploratory analysis"
---

## Purpose

Exploration is where students learn the shape of the data before fitting formal models. This section should summarize distributions, identify outliers, and reveal possible relationships worth modeling.

```{r}
#| eval: false
library(tidyverse)
source("../R/analysis_functions.R")
```

## Example summaries

```{r}
#| eval: false
surveys_df |>
  count(habitat_type)

surveys_df |>
  summarize(
    mean_bees = mean(bee_count, na.rm = TRUE),
    sd_bees = sd(bee_count, na.rm = TRUE),
    mean_temp = mean(temperature, na.rm = TRUE)
  )
```

## Example exploratory plots

```{r}
#| eval: false
surveys_df |>
  ggplot(aes(x = habitat_type, y = bee_count)) +
  geom_boxplot()
```

```{r}
#| eval: false
plot_temporal_patterns(phenology_df)
```

## Questions to ask during exploration

- Are counts heavily zero-inflated?
- Do some habitats show consistently different abundance?
- Are temperature and bee visits associated?
- Are there seasonal peaks by elevation band?
- Are there obvious site-level clusters?

## Transition to modeling

Exploration should motivate the modeling chapter. By the end of this document, the reader should understand why particular predictors and model forms were chosen.
