---
title: "03: Statistical modeling"
---

## Purpose

This section contains the formal models used to answer the project questions. It is usually the most thesis-like part of the repository because it ties biological questions to specific statistical machinery.

```{r}
#| eval: false
source("../R/analysis_functions.R")
```

## Abundance model

The example repository includes a generalized linear mixed model for bee abundance using elevation, slope, transformed aspect, habitat, temperature, and wind speed, with random intercepts for site and date.

```{r}
#| eval: false
mod_abundance <- abundance_model(surveys_df)
summary(mod_abundance)
```

## Habitat comparison

The repository also includes a habitat-level aggregation followed by ANOVA.

```{r}
#| eval: false
habitat_results <- habitat_analysis(surveys_df)
summary(habitat_results$model)
```

## Phenological analysis

Temporal matching between flowering and bee activity can be modeled with a GAM.

```{r}
#| eval: false
mod_pheno <- phenology_analysis(phenology_df)
summary(mod_pheno)
```

## Spatial analysis

At the site level, students can test for spatial autocorrelation and fit a distribution model.

```{r}
#| eval: false
moran_out <- spatial_autocorr(spatial_df)
moran_out

sdm_out <- distribution_model(spatial_df)
summary(sdm_out)
```

## What students should explain

Students should not stop at output. They should state:

- the response variable,
- the predictors,
- why the model family is appropriate,
- what random effects represent,
- and how the model answers the biological question.
