---
title: "04: Publication figures"
---

## Purpose

This chapter turns exploratory and modeling output into polished figures saved to `figs/` for use in a thesis, defense, or manuscript.

```{r}
#| eval: false
library(ggplot2)
source("../R/analysis_functions.R")
```

## Temporal figure

```{r}
#| eval: false
p_time <- plot_temporal_patterns(phenology_df)
p_time

ggsave(
  filename = "../figs/temporal-patterns.png",
  plot = p_time,
  width = 10,
  height = 7,
  dpi = 300
)
```

## Spatial figure

```{r}
#| eval: false
p_space <- plot_spatial_distribution(spatial_df)
p_space

ggsave(
  filename = "../figs/spatial-distribution.png",
  plot = p_space,
  width = 10,
  height = 7,
  dpi = 300
)
```

## Figure standards

A thesis repository should save figures rather than rely on temporary plots in the R session. That makes it easier to:

- reuse figures in writing,
- track changes over time,
- and regenerate publication-ready output.

## Final note

Students should label axes carefully, use readable themes, and choose figure dimensions that fit the intended document.
